---
title: "De-bugging exercises: Data cleaning & management"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(dplyr)
library(magrittr)
library(bst290)

data(ess)
knitr::opts_chunk$set(echo = FALSE)
```

# Data cleaning & management 

## De-bugging exercises {data-progressive=TRUE}

Here are again some exercises that require you to fix problems with code chunks. Hint: Look for typos, missing or misplaced symbols, and check the parentheses.

### Problem No. 1

In this scenario, you want to select three variables from the `ess` dataset, but all you get is an error message:
```{r ex1, exercise = TRUE}
ess %>% 
  select(idno,gndr,agea
```

<div id="ex1-hint">
**Hint:** Everything that is opened must also be closed...
</div>

### Problem No. 1: Solution

The last closing parenthesis was of course missing.

```{r ex1-a, eval=T, echo=T}
ess %>% 
  select(idno,gndr,agea)
```

### Problem No. 2

Now you want to link two operations, where the first selects variables and the second filters observations. But, again, something does not work:
```{r ex2, exercise = TRUE}
ess %>% 
  select(idno,cntry,agea,vote,stflife)
  filter(agea>=25)

```

<div id="ex2-hint">
**Hint:** Check if there are any missing symbols or operators.
</div>

### Problem No. 2: Solution

The problem was that the pipe operator was missing at the end of the `select()`-line, and `R` did not understand that the operations should be linked together:
```{r ex2-a, eval=T, echo=T}
ess %>% 
  select(idno,cntry,agea,vote,stflife) %>% 
  filter(agea>=25)

```

### Problem No. 3

Here you want to filter the data so that only women are left (where `gndr` is equal to `Female`), but somehow `R` refuses to do what you want:
```{r ex3, exercise = TRUE}
ess %>% 
  filter(gndr="Female")
```

<div id="ex3-hint">
**Hint:** How do you tell `R` that two things *should* be equal?
</div>

### Problem No. 3: Solution

The problem was with the equal signs: If you want tell `R` that something *should* or *has to be* equal to something else, you always use double equal signs (`==`).

```{r ex3-a, eval=T, echo=T}
ess %>% 
  filter(gndr=="Female")
```

### Problem No. 4

OK, this one might be a bit mean --- but maybe you can figure it out anyways? 

You want to select some variables from the `ess` dataset, this time without using the pipe but by specifying the dataset explicitly in the `select()` command:
```{r ex4, exercise = TRUE}
select(data = ess,
       agea,idno,gndr)
```

<div id="ex4-hint">
**Hint:** Take a close look at the `data=ess` part --- there was a special thing about `select()` and `filter()` where you had to use a specific punctuation symbol...
</div>

### Problem No. 4: Solution

The problem was the special way you name the dataset you want to use in `select()` (or also `filter()`): You have to add a dot before `data` so that it reads `.data = ess`, not `data = ess`:

```{r ex4-a, eval=F, echo=T}
select(.data = ess,
       agea,idno,gndr)
```

A more convenient alternative is to use the pipe operator from the start:
```{r ex4-b, eval=T, echo=T}
ess %>% 
  select(agea,idno,gndr)
```

### Conclusion

**That is it for this part!** Next: Data visualization!
