---
title: "De-bugging exercises: Getting to know your data"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(bst290)
knitr::opts_chunk$set(echo = FALSE)
```

# Getting to know your data

## De-bugging exercises {data-progressive=TRUE}

As before: De-bugging is unfortunately a part of working with computer code. In the following exercises, you get a piece of code that does not work and have to figure out how to fix it.

In all exercises, you will work with the small `ess` dataset that you used in the main tutorial and which you should therefore be familiar with (otherwise, quickly go over the main tutorial again to refresh your memory).

### Problem No. 1

Here you want to calculate the average age of your respondents --- but `R` refuses...
```{r ex1, exercise=T}
average(ess$agea)
```

<div id="ex1-hint">
**Hint:** Was the relevant function really called `average()`? There was also this other word for 'average'...
</div>

### Problem No. 1: Solution

Obviously, the problem was that the function to calculate an average in `R` is called `mean()`, and not `average()`. You may find that strange or stupid, but it just is that way...

Here is how you make that code work:
```{r ex1-a, echo=T, eval=F}
mean(ess$agea)
```

<!-- ### Problem No. 2 -->

<!-- Now you want to calculate the average level of trust in other people (`ppltrst`). You have the correct function, `mean()`, but you get `NA` as a result. Disappointing. What could be the problem here? -->
<!-- ```{r ex2, exercise = T} -->
<!-- mean(ess$ppltrst) -->
<!-- ``` -->

<!-- <div id="ex2-hint"> -->
<!-- **Hint:** Use the `oppsumtabell()` function to see if there are any missings (`NA`s) in the data (`oppsumtabell(dataset = ess, variables = "ppltrst")`; you can run this right here in the little console). Can this be the cause of your problem? How could you "remove" the problematic observations? -->
<!-- </div> -->

<!-- ### Problem No. 2: Solution -->

<!-- The problem was that there was one missing observation (`NA`s) in the `ppltrst` variable. You can see this if you run the `oppsumtabell()` function, listed as `Missings` all the way at the bottom: -->
<!-- ```{r ex2-a1, echo=T, eval=T} -->
<!-- oppsumtabell(dataset = ess,  -->
<!--              variables = "ppltrst") -->
<!-- ``` -->

<!-- If you now remove this one missing observation from the data before calculating the mean, the problem is fixed. All you need to do is to set the `na.rm` option to `TRUE`: -->
<!-- ```{r ex2-a2, echo=T, eval=T} -->
<!-- mean(ess$ppltrst, -->
<!--      na.rm = TRUE) -->
<!-- ``` -->

<!-- You also see that the result from the `mean()` function corresponds to the number you get from `oppsumtabell()` (under `Average`) -- this is because `oppsumtabell()` automatically removes any missing observations from the data. -->

### Problem No. 2

You want to see the first 20 observations of the `ess` dataset -- but `R` only shows you the first six. How can you get 20?
```{r ex3, exercise = T}
head(ess)
```

<div id="ex3-hint">
**Hint:** There is an option for the `head()` function that lets you decide how many observations are printed. `n=`...
</div>

### Problem No. 2: Solution

This one is solved by setting the number of rows to 20 with the `n` option:
```{r ex3-a, eval=T, echo=T}
head(ess,
     n = 20)

```

### Problem No. 3

You want to calculate the standard deviation of the `agea` variable, and you definitely have the correct function and variable --- but something still causes an error:

```{r ex4, exercise = T}
sd(ess@agea)
```

<div id="ex4-hint">
**Hint:** Which symbol do you need to use to select a specific variable from a dataset?
</div>

### Problem No. 3: Solution

The issue was that the wrong symbol was used to indicate that `R` should use a specific variable from the `ess` dataset: This is done with the `$` symbol, not with `@`:
```{r ex4-a, eval=T, echo=T}
sd(ess$agea)
```

### A last quiz

```{r quiz, echo=F}
quiz(caption = "Final (brief) quiz",
     question("What is the correct function to get a list of the distinct categories of a variable?",
              answer("`unique()`", correct = T),
              answer("`distinct()`", message = "Not quite - try again!"),
              answer("`categories()`", message = "Unfortunately not - but don't give up, you know this!"),
              random_answer_order = T,
              allow_retry = T),
     question("How does `R` like to store categorical variables?",
              answer("As *factors*", correct = T),
              answer("As *categoricals*", message = "Not quite - try again!"),
              answer("As *levels*", message = "Unfortunately not - but don't give up, you know this!"),
              random_answer_order = T,
              allow_retry = T),
     question("How can you get the last few observations of a dataset?",
              answer("`tail()`", correct = T),
              answer("`end()`", message = "Not quite - try again!"),
              answer("`last()`", message = "Almost...keep trying."),
              random_answer_order = T,
              allow_retry = T))

```

### Conclusion

**Congrats, you are done!** Now you know how you can get summary statistics and familiarize yourself with a new dataset.

