---
title: "De-bugging exercises: Getting your data into R"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(haven)
library(readxl)
knitr::opts_chunk$set(echo = FALSE)
```

# Getting your data into R

## De-bugging exercises {data-progressive=TRUE}

When you work with `R` (or any other programming language) you will sooner or later run into "bugs" --- problems with your code that cause error messages and which you need to fix to be able to move on. The process of fixing "bugs" is called "de-bugging".

Error messages can seem quite dramatic when you have not much experience with programming and coding, and it can take a bit to solve any individual "bug". Over time, however, you will learn one very important lesson: **Most "bugs" are simply the result of typos**: You forgot to put a comma where `R` expects a comma to be; you misspelled something; you mixed up commands; you forgot to put text in quotation marks. Simple (and, sometimes, embarrassing) mistakes.

This is the first of several exercises that allow you to practice and gain confidence with "de-bugging". Your progress is always saved, but you can also start over by clicking "Start over" on the left.

You will get little chunks of `R` code that have a "bug" hidden somewhere --- it is your job to "de-bug" the code and get it to work. **You can test if the code works by hitting the "Run" button on the right.**

(Hint: Look for typos, little mix-ups, missing quotation marks, or parentheses.)

### Problem No. 1

In this first problem, we want to download and read the Stata (.dta) version of the *Comparative Political Data Set*, but the code does not seem to work. Can you help?
```{r ex1, exercise=T}
cpds <- read_sav("https://www.cpds-data.org/images/Update2020/CPDS_1960-2018_Update_2020.dta")

# This prints the first few observations of the dataset
head(cpds)
```

<div id="ex1-hint">
**Hint:** Check if the function ("command") is correct for the type of dataset file (.dta) you want to read.
</div>

### Problem No. 1: Solution

The problem: The wrong `read_` function was used. To read a Stata dataset, we need to use the `read_dta()` function, not the `read_sav()` function. This is the correct solution:
```{r ex1-a, echo=T, eval=F}
cpds <- read_dta("https://www.cpds-data.org/images/Update2020/CPDS_1960-2018_Update_2020.dta")

# This prints the first few observations of the dataset
head(cpds)
```

Recall also what the `head()` function does: It shows you the first few observations in your dataset.

### Problem No. 2

Here is the second problem. Can you fix the bug?
```{r ex2, exercise=T}
mepv <- read_sav(http://www.systemicpeace.org/inscr/MEPVv2018.sav)

# This prints the first few observations of the dataset
head(mepv)
```

<div id="ex2-hint">
**Hint:** A link is text, and text always goes between ... marks.
</div>

### Problem No. 2: Solution

Could you figure out what the problem was? Again one of the most common ones: Missing quotation marks! We have to tell R not to treat the link as a command but to treat it simply like a piece of text.

This is the correct solution
```{r ex2-a, echo=T, eval=F}
mepv <- read_sav("http://www.systemicpeace.org/inscr/MEPVv2018.sav")

# This prints the first few observations of the dataset
head(mepv)
```

### Problem No. 3

This "bug" can occur anytime, whether you download and read data or work with objects that are already assigned and stored. Do you see where the problem is?

```{r ex3, exercise = T}
read_dta("https://www.cpds-data.org/images/Update2020/CPDS_1960-2018_Update_2020.dta")

# This prints the first few observations of the dataset
head(cpds)
```

<div id="ex3-hint">
**Hint:** Is the dataset assigned to `cpds`?
</div>

### Problem No. 3: Solution

This one might have been a bit tricky --- the problem was that the dataset was downloaded but not stored as an object in `R`. You always need to store data, values,... in objects using the assignment operator (`<-`) if you want to use them later:
```{r ex3-a, eval=F, echo=T}
cpds <- read_dta("https://www.cpds-data.org/images/Update2020/CPDS_1960-2018_Update_2020.dta")

# This prints the first few observations of the dataset
head(cpds)
```

### A quick quiz to wrap up

You made it through the "de-bugging" exercises. If you solved all of them, great job! If you did not manage, do not worry --- this just takes a bit of practice.

Before you finish, can you answer this little quiz?

```{r quiz, echo=F}
quiz(caption = "Final quiz",
question("Which package does NOT allow you to directly download datasets from the internet?",
         answer("`haven`", message = "Not quite. Try again!"),
         answer("`readxl`", correct = T,
                message = "Correct! When you use `readxl` to load Excel dataset files, you first need to save them on your computer before you can load them into `R`."),
         allow_retry = TRUE,
  random_answer_order = TRUE),
question("Now a question on how you work with PACKAGES: Which command ('function') do you use to INSTALL packages?",
         answer("`install()`", message = "Almost, keep trying!"),
         answer("`load.packages()`", message = "Not quite - you can do this!"),
         answer("`install.packages()`", message = "Correct! You install packages with `install.packages()`, for example: `install.packages('haven')`.", correct = T),
         allow_retry = TRUE,
  random_answer_order = TRUE),
question("Finally, how do you LOAD a package after you installed it?",
         answer("`load()`", message = "Not quite - you can do this!"),
         answer("`load.packages()`", message = "Almost, keep trying!"),
         answer("`library()`", message = "Correct. You use the library() function, for example: `library(haven)`.",
                correct = TRUE),
         allow_retry = TRUE,
  random_answer_order = TRUE)
)
```

### Conclusion

You made it! I hope you see now that error messages --- of which we all get many --- are usually trivial and easily fixed.

Ready to move on to the next tutorial on "Getting to know your data"?

