---
title: "De-bugging exercises: Data visualization"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(ggplot2)
library(bst290)
library(dplyr)
library(magrittr)

data(ess)

knitr::opts_chunk$set(echo = FALSE)
```

# Data visualization

## De-bugging exercises {data-progressive=TRUE}

You should know the drill by now: You get a few code chunks, all of them are supposed to produce `ggplot2` graphs but all of them have some problem(s) in them that generate only error messages. Your job is to fix the bugs and make the code work properly.

### Problem No. 1

Here the task is to create a scatterplot with `geom_point()`, but the points are not shown. This should not be hard to figure out, however:
```{r ex1, exercise = T}
ggplot(data = ess, aes(x = height, y = weight))
  geom_point()
```

<div id="ex1-hint">
**Hint:** You need a particular mathematical symbol to add new layers to `ggplot2` graphs.
</div>

### Problem No. 1: Solution

The problem was a lacking plus sign (`+`) after the first line. Without the `+`, `R` does not understand that the two functions (`ggplot()` & `geom_point()`) belong together and therefore gets confused. 

Once you add the `+`, everything works:
```{r ex1-a, eval=T, echo=T}
ggplot(data = ess, aes(x = height, y = weight)) +
  geom_point()
```

### Problem No. 2

You are still working with the same scatterplot, and now you are trying to make the graph publication-ready by adding informative labels. Unfortunately, `R` does not seem to like the code that it gets here. Can you find and fix this bug?

```{r ex2, exercise = T}
ggplot(data = ess, aes(x = height, y = weight)) +
  geom_point() +
  xlab(Body height (cm)) +
  ylab(Body weight (kg)) +
  theme_bw()
```

<div id="ex2-hint">
**Hint:** When you add axis labels, you add text --- and text always goes between...what?
</div>

### Problem No 2: Solution

The issue: The text you wanted to add with `xlab()` and `ylab()` was not put in quotation marks. Therefore, `R` thought that *Body height (cm)* and *Body weight (kg)* would refer to some functions it is supposed to run --- but which obviously do not exist. 

Once you put the text in quotation marks, the graph displays as intended:
```{r ex2-a, echo=T, eval=T}
ggplot(data = ess, aes(x = height, y = weight)) +
  geom_point() +
  xlab("Body height (cm)") +
  ylab("Body weight (kg)") +
  theme_bw()
```

### Problem No. 3

Here the task is to create a different type of graph, a bar graph showing the number of male and female respondents in the `ess` dataset --- but the right `geom_` still needs to be specified.

```{r ex3, exercise = T}
ggplot(data = ess, aes(x = gndr))
```

<div id="ex3-hint">
**Hint:** Which geometric object or 'geom' produces a bar graph?
</div>

### Problem No. 3: Solution

You obviously needed to add `geom_bar()` as an additional graph layer. Since you want it to represent the overall numbers --- or 'counts' --- of male and female respondents, you could also be explicit about it and add `stat = "count"` --- but it works also without:
```{r ex3-a, echo=T, eval=T}
ggplot(data = ess, aes(x = gndr)) +
    geom_bar(stat = "count")
```

### Problem No 4.

To solve this last problem, you need to combine your data cleaning and management skills with your knowledge about data visualization.

You want to create a *bar graph* that shows the average ages of respondents, depending on whether or not they were born in Norway. The relevant variables are:

* `agea`, which measures respondents' ages;
* `brncntr`, which records whether or not respondents were born in Norway;

**Heads up:** This one might take a while to solve...
```{r ex4, exercise = T}
ess %>% 
  group_by() %>% 
  summarize(mean_age = ) %>% 
  ggplot(aes(x = , y = mean_age)) +
    geom_
```

<div id="ex4-hint">
**Hint:** You still need to add quite a bit of code. Have a look at Section 4 of the Tutorial.
</div>

### Problem No 4: Solution

Could you figure it out? If yes, congrats, this one was not that beginner-friendly!

You needed to add:

* Which variable you wanted to `group_by()`: `brncntr`
* How you wanted to `summarize()` the `agea` variable: `mean_age = mean(agea, na.rm = T)`
* Which geometric object you wanted to have in the bar graph: `geom_bar()`, obviously, and...
* What the bars are supposed to show: Here, the 'raw' data, which means `stat = "identity"`

Everything together should look like this:
```{r ex4-a, eval=T, echo=T}
ess %>% 
  group_by(brncntr) %>% 
  summarize(mean_age = mean(agea, na.rm = T)) %>% 
  ggplot(aes(x = brncntr, y = mean_age)) +
    geom_bar(stat = "identity")
```

### Conclusion

**Great job, you are done with these de-bugging exercises**. Now you know the fundamentals of visualizing data with `ggplot2` in `R`!

If you like, you can still polish the last graph a bit to make it publication-ready --- otherwise, get ready for the next part of this course: *Statistical inference*.
```{r ex5, exercise = T}
ess %>% 
  group_by(brncntr) %>% 
  summarize(mean_age = mean(agea, na.rm = T)) %>% 
  ggplot(aes(x = brncntr, y = mean_age)) +
    geom_bar(stat = "identity")
```
